---
title: "Influence of Summer Temperature Rise On Grepes Harvest Dates in Bordeaux, France "
output: rmarkdown::github_document
---
by Usman Ahmed

## Project Description

In northern countries, the grapevine typically breaks bud in April-May, flowers occur in June, and the veraison at the end of July or during August. The ripening phase initiates at the veraison and finishes when grapes have reached a certain level of maturity. Currently, in France, the grape harvest is most often scheduled between 10 September and 10 October (Van Leeuwen et al., 2008). Due to strong control over temperature on grape ripening, The grapes harvest dates (GHD) are closely related to growing season temperature (GST). GHD vlaues are following the nomenclature adopted by the Angot, 1883 and are expressed as the number of days after the 31 August (e.g. 1st September is 1, 1st October is 31, 29th August is -2, etc.)
GST is the mean of monthly mean of daily maximum temperatures in &deg;C (Harris et al., 2014) from April to August.

```{r warning=FALSE, message=FALSE}
library(date)
library(ggplot2)
library(foreign)
```

Importing data

```{r }
GHD<- "raw_data/Project_4_GHD.csv"

GHD <- read.csv(GHD, header=T, skip=0, sep=";", na.strings="NA")

Bor_tmax<- "raw_data/Project_4_Bor.tmax.csv"
Bor_tmax <- read.csv(Bor_tmax, header=T, skip=0, sep=";", na.strings="NA")
```

Start GHD from 1901 to synchronize the time series with GST, since GST starts from 1901

```{r}
GHD <- GHD [!(GHD$Year.GHD<1901),]
```

### The regions for which GHDs are available, are Bordeaux, Burgundy, Champagne 1, Low Loire Valley, Spain and Switzerland

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.align='center', fig.height=5, fig.width=9}
ggplot(GHD, aes(x=GHD$Year.GHD))+
  geom_line(aes(y=GHD$Bor, colour="Bor"), size=0.8)+
  geom_line(aes(y=GHD$Bur, colour="Bur"), size=0.8)+
  geom_line(aes(y=GHD$Cha2, colour="Cha2"), size=0.8)+
  geom_line(aes(y=GHD$LLV, colour="LLV"), size=0.8)+
  geom_line(aes(y=GHD$Spa, colour="Spa"), size=0.8)+
  geom_line(aes(y=GHD$Swi, colour="Swi"), size=0.8)+
  
  
  
  scale_color_manual("Legend", breaks=c("Bor","Bur","Cha2","LLV","Spa","Swi"), values=c("red", "green","blue","black","brown","Yellow"))+
  
  coord_cartesian(xlim = c(1900,2015), ylim = c(65, -10))+
  scale_y_reverse( lim=c(65,-10))+
  
  labs(x="Year", y="GHD")+ ggtitle("Grape Harvest Dates (GHD)")+
  theme_bw()+
  theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),
        plot.title = element_text(hjust = 0.5))
```

## GST(mean of tmax from Apri-Aug) and GHD of Bordeaux

An increasing trend of mean summer temperature coincides early grapes harvest dates at Bordeaus region.

```{r fig.align='center', fig.height=5, fig.width=9}
year <- Bor_tmax$Year
Bor_gst <- rowMeans(Bor_tmax[,c(5:9)], na.rm=TRUE)

GST <- data.frame(year, Bor_gst)
GST <- GST [!(GST$year>2010),] 

par(mar=c(5,5,3,5)) 
plot(GHD$Year.GHD, GHD$Bor, type ="l",col="green", xlim = c(1900,2010), ylim = c(55,-5),
     ylab="Bordeaux GHD", xlab="Year")
par(new=TRUE) 
plot(GST$year, GST$Bor_gst, type ="l", col="red",
     xlim = c(1900,2010), ylim = c(20,30),
     axes= FALSE, xlab = NA, ylab = NA) 
axis(side = 4, col="red")
mtext(side = 4, line = 3, 'Bordeaux GST (°C)')
title(main = list("GHD-GST (April-August) of Bordeaux Region",col = "black", font = 1))
```

```{r fig.align='center', fig.height=5, fig.width=9}
plot(GST$Bor_gst, GHD$Bor, type="p", col = "green", ylab="Bordeaux GHD", xlab="Bordeaux GST (°C)")
abline(mod<-lm(GHD$Bor~GST$Bor_gst),col="red",lwd=1.5)
corr <- cor(GST$Bor_gst, GHD$Bor,  method = "pearson", use = "complete.obs")
legend(x='topright', 
       legend=paste('Corr. Coef. =',round(cor(GST$Bor_gst, GHD$Bor,  method = "pearson", use = "complete.obs"),3)))
title(main = list("GHD-GST Correlation of Bordeaux Region",col = "black", font = 1)) 
```

High negative correlation shows that increasing summer mean temperature is resulting earlier grapes harvest in the Bordeaux region. A mean summer temperatue of >24&deg;C will result early harvest in August or in early September. Similar trends of early GHD are also observed for other grapes harvest regions. 

*Data Source:University of Bremen